---
title: "iTRAQI"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(gstat)
library(tidyverse)
library(sf)
library(rnaturalearth)
library(ggspatial)
library(progress)
```

#### resources used:
* kriging https://rpubs.com/nabilabd/118172
* plots https://r-spatial.github.io/sf/articles/sf5.html
* understanding crs https://www.earthdatascience.org/courses/earth-analytics/spatial-data-r/understand-epsg-wkt-and-other-crs-definition-file-types/
* changing crs https://stackoverflow.com/questions/50372533/changing-crs-of-a-sf-object
* spatial aggregations https://cengel.github.io/R-spatial/spatialops.html

# load inputs
```{r}
df_times <- read.csv("input/QLD_locations_with_RSQ_times_20220210.csv")
coordinates(df_times) <- ~ x + y

# qld_bounary <- st_read("input/qld_state_polygon_shp/QLD_STATE_POLYGON_shp.shp")
qld_bounary <- read_sf("input/qld_state_polygon_shp/QLD_STATE_POLYGON_shp.shp")


qld_SAs <- st_read("input/qld_sa_zones/MB_2016_QLD.shp")
# ggplot(data = world) +
#   geom_sf() +
#   geom_sf(data=qld_SAs, aes(fill=SA3_NAME16), col = NA) +
#   coord_sf(xlim = c(100.00, 160.00), ylim = c(-45.00, -10.00), expand = T) + 
#   guides(fill="none")

```

# base map and points
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_bounary, color="blue") +
  labs( x = "Longitude", y = "Latitude") +
  annotation_scale(location = "bl", width_hint = 0.5) +
  geom_sf(data=qld_SAs, aes(fill=SA3_NAME16), col = NA) +
  geom_point(data=as.data.frame(df_times), aes(x,y))+ 
  guides(fill="none") +
  coord_sf(xlim = c(100.00, 160.00), ylim = c(-45.00, -10.00), expand = T)

```

# kriging - select variogram model (Gaussian looks best)
```{r}
lzn_vgm <- variogram(acute_time~1, df_times)
lzn_fit <- fit.variogram(lzn_vgm, model=vgm("Gau"))

plot(lzn_vgm, lzn_fit)
# show.vgms()
```

# kriging - create spatial domain to interpolate over
```{r}
map <- read_sf("input/qld_state_polygon_shp/QLD_STATE_POLYGON_shp.shp")
aus <- raster::getData('GADM', country = 'AUS', level = 1)
grid <- makegrid(aus[aus$NAME_1 == "Queensland",], cellsize = 0.05)
pnts_sf <- st_as_sf(grid, coords = c('x1', 'x2'), crs = st_crs(map))

pnts <- pnts_sf %>% mutate(
  intersection = as.integer(st_intersects(geometry, map))
) %>%
  filter(!is.na(intersection)) %>%
  st_coordinates() %>% 
  as.data.frame()
coordinates(pnts) <- ~ X + Y

```

# kriging - generate interpolations
```{r}
lzn_kriged <- krige(acute_time ~ 1, df_times, pnts, model=lzn_fit)%>%
  as.data.frame

lzn_kriged %>%
  ggplot(aes(X, Y)) + 
  geom_tile(aes(fill=var1.pred)) + 
  coord_equal()+
  scale_fill_gradient(low = "yellow", high="red")

```

# add interpolation layer to map
```{r}
ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_bounary, color="blue") +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(100.00, 160.00), ylim = c(-45.00, -10.00), expand = T) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  geom_tile(data=lzn_kriged, aes(X, Y, fill=var1.pred)) +
  geom_point(data=as.data.frame(df_times), aes(x,y)) +
  scale_fill_gradient(low = "yellow", high="red")

```

# Join the SA2s polygons together: https://gis.stackexchange.com/questions/63577/joining-polygons-in-r
### This isn't reducing the plotting time as much as I had thought. I think the combined polygons (after st_union()) still contain every polygon that's been aggregated rather than keeping the largest polygon that contains them all (outside boarder). Not sure how to do this but there may be some better way out there.
```{r}

do_union <- function(x){
  pb$tick()
  st_union(x, is_coverage = TRUE)
}

aggregate_by_SA <- function(qld_sf, SA_number){
  sa_main <- glue::glue('SA{SA_number}_{ifelse(SA_number %in% c(3,4), "CODE", "MAIN")}16')
  message(glue::glue('----- grouping polygons within SA{SA_number} -----'))
  num_ticks <- length(unique(qld_sf[[sa_main]]))
  pb <<- progress_bar$new(
    format = "[:bar] :current/:total (:percent) elapsed :elapsed eta :eta", 
    total=num_ticks
  )
  
  qld_sf %>%
    # https://gis.stackexchange.com/questions/321281/using-sf-to-combine-polygons-that-share-borders
    group_by(across(all_of(sa_main))) %>%
    summarize(geometry=do_union(geometry))
}

qld_SA2s <- aggregate_by_SA(qld_sf=qld_SAs, SA_number=2)
qld_SA1s <- aggregate_by_SA(qld_sf=qld_SAs, SA_number=1)

ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_SA2s, aes(fill=1)) +
  coord_sf(xlim = c(138, 153.00), ylim = c(-30.00, -10.00), expand = T) +
  guides(fill="none") +
  scale_fill_gradient(low = "yellow", high="red")
```

# find out why there is a mix of polygons and multipolygons... 
##### spoiler: multipolygons where there are islands
```{r}
mlt_idx <- c()
for(i in 1:nrow(qld_SA2s)){
  geo <- qld_SA2s$geometry[i]
  cl <- class(geo[[1]])[2]
  if(cl=="MULTIPOLYGON"){
    mlt_idx <- append(mlt_idx, i)
  }
}

ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_SA2s[mlt_idx,], aes(fill=1)) +
  geom_sf(data=qld_SA2s[-mlt_idx,], aes(fill=0)) +
  coord_sf(xlim = c(138, 153.00), ylim = c(-30.00, -10.00), expand = T) +
  guides(fill="none") +
  scale_fill_gradient(low = "yellow", high="red")
```

# spatial join from SA2 polygons to interpolated values
```{r}
coordinates(lzn_kriged) <- ~ X + Y
lzn_kriged_sf <- st_as_sf(lzn_kriged)

lzn_kriged_sf <- st_set_crs(lzn_kriged_sf, 4326)
lzn_kriged_sf <- st_transform(lzn_kriged_sf, crs = 4326)

qld_SA2s <- st_transform(qld_SA2s, crs = 4326)
qld_SA2s_with_int_times <- st_join(qld_SA2s, lzn_kriged_sf)
# https://ryanpeek.org/2019-04-29-spatial-joins-in-r/

qld_SA2s_agged_times <- 
  qld_SA2s_with_int_times %>% 
  na.omit() %>%
  group_by(SA2_MAIN16) %>%
  mutate(mean=mean(var1.pred, na.omit=TRUE)) %>% 
  ungroup()

ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_SA2s_agged_times, aes(fill=mean), col = NA) +
  coord_sf(xlim = c(138, 153.00), ylim = c(-30.00, -10.00), expand = T) +
  guides(fill="none") +
  scale_fill_gradient(low = "yellow", high="red")

```

# spatial join from SA1 polygons to interpolated values
```{r}
qld_SA1s <- st_transform(qld_SA1s, crs = 4326)
qld_SA1s_with_int_times <- st_join(qld_SA1s, lzn_kriged_sf)
# https://ryanpeek.org/2019-04-29-spatial-joins-in-r/

qld_SA1s_agged_times <- 
  qld_SA1s_with_int_times %>% 
  na.omit() %>%
  group_by(SA1_MAIN16) %>%
  mutate(mean=mean(var1.pred, na.omit=TRUE)) %>% 
  ungroup()

ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_SA1s_agged_times, aes(fill=mean), col = NA) +
  coord_sf(xlim = c(138, 153.00), ylim = c(-30.00, -10.00), expand = T) +
  guides(fill="none") +
  scale_fill_gradient(low = "yellow", high="red")

```

# write interpolation layers to disk
```{r}
saveRDS(qld_SA1s_agged_times, "input/layers/SA1s_acute.rds")
saveRDS(qld_SA2s_agged_times, "input/layers/SA2s_acute.rds")
```