---
title: "iTRAQI"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(sp)
library(gstat)
library(tidyverse)
library(sf)
library(rnaturalearth)
library(ggspatial)
library(progress)
library(leaflet)
library(leaflet.extras)
```

#### resources used:
* kriging https://rpubs.com/nabilabd/118172
* plots https://r-spatial.github.io/sf/articles/sf5.html
* understanding crs https://www.earthdatascience.org/courses/earth-analytics/spatial-data-r/understand-epsg-wkt-and-other-crs-definition-file-types/
* changing crs https://stackoverflow.com/questions/50372533/changing-crs-of-a-sf-object
* spatial aggregations https://cengel.github.io/R-spatial/spatialops.html

# load inputs
```{r}
df_times <- read.csv("input/QLD_locations_with_RSQ_times_20220210.csv")
coordinates(df_times) <- ~ x + y

qld_bounary <- read_sf("input/qld_state_polygon_shp/QLD_STATE_POLYGON_shp.shp")
world <- ne_countries(scale = "medium", returnclass = "sf")
qld_SAs <- st_read("input/qld_sa_zones/MB_2016_QLD.shp")
# ggplot(data = world) +
#   geom_sf() +
#   geom_sf(data=qld_SAs, aes(fill=SA3_NAME16), col = NA) +
#   coord_sf(xlim = c(100.00, 160.00), ylim = c(-45.00, -10.00), expand = T) + 
#   guides(fill="none")

```

# base map and points
```{r}
ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_bounary, color="blue") +
  labs( x = "Longitude", y = "Latitude") +
  annotation_scale(location = "bl", width_hint = 0.5) +
  geom_sf(data=qld_SAs, aes(fill=SA3_NAME16), col = NA) +
  geom_point(data=as.data.frame(df_times), aes(x,y))+ 
  guides(fill="none") +
  coord_sf(xlim = c(100.00, 160.00), ylim = c(-45.00, -10.00), expand = T)

```



# kriging - create spatial domain to interpolate over
```{r}
aus <- raster::getData('GADM', country = 'AUS', level = 1)
grid <- makegrid(aus[aus$NAME_1 == "Queensland",], cellsize = 0.1)
pnts_sf <- st_as_sf(grid, coords = c('x1', 'x2'), crs = st_crs(qld_bounary))

pnts <- pnts_sf %>% mutate(
  # https://gis.stackexchange.com/a/343479
  intersection = as.integer(st_intersects(geometry, qld_bounary))
) %>%
  filter(!is.na(intersection)) %>%
  st_coordinates() %>% 
  as.data.frame()

# add centroids of all polygons to pnts to ensure there's at least one interpolated value within
centroids <- 
  qld_SAs %>%
  st_centroid() %>%
  st_coordinates() %>%
  na.omit() %>% 
  as.data.frame()

pnts2 <- rbind(pnts, centroids)
coordinates(pnts) <- ~ X + Y
coordinates(pnts2) <- ~ X + Y

```

# kriging - select variogram model 
### Gaussian looks best but Spherical matches expectation
```{r}
# "Gau" model gives better model fit but does not match up with our 
# interpolations in ArcGIS like the Spherical ("Sph") option does
vgm_model <- "Sph" # Exc or Cir looks best for acute

# acute variogram
lzn_vgm <- variogram(acute_time~1, df_times)
lzn_fit_acute <- fit.variogram(lzn_vgm, model=vgm(vgm_model))
plot(lzn_vgm, lzn_fit_acute)

vgm_model <- "Sph" # Sph or Cir looks best for rehab (most sensible interpolations)
# rehab variogram
lzn_vgm <- variogram(rehab_time~1, df_times)
lzn_fit_rehab <- fit.variogram(lzn_vgm, model=vgm(vgm_model))
plot(lzn_vgm, lzn_fit_rehab)

show.vgms()
```

```{r, echo=F, results='hide'}
f_rehab_vgm_model_fit <- function(model){
  lzn_vgm <- variogram(rehab_time~1, df_times)
  lzn_fit_rehab <- fit.variogram(lzn_vgm, model=vgm(model))
  plot(lzn_vgm, lzn_fit_rehab, main=glue::glue("using a {model} model"))
}

# map(c("Sph", "Exc", "Ste", "Cir", "Lin", "Pen"), f_rehab_vgm_model_fit)
# f_rehab_vgm_model_fit("Cir")
```

```{r, echo=F, results='hide'}
# kriging cross-validation
vgm_model <- "Exp"

# lzn_vgm <- variogram(acute_time~1, df_times)
# lzn_fit_acute <- fit.variogram(lzn_vgm, model=vgm(vgm_model))

f_krige_cv <- function(model, time, nfold=10){
  print(model)
  m <- vgm(.59, model, 874, .04)
  # five-fold cross validation:
  if(time=="acute"){
    x <- krige.cv(acute_time~1, df_times, m, nmax = 40, nfold=nfold)
  }
  if(time=="rehab"){
    x <- krige.cv(rehab_time~1, df_times, m, nmax = 40, nfold=nfold)
  }
  bubble(x, "residual", main = glue::glue("{time} time: 5-fold CV residuals with {model} model"))
}

# map(c( "Exp", "Sph", "Gau", "Exc", "Mat", "Ste", "Cir", "Lin", "Bes", "Pen", "Per", "Wav", "Hol"), function(x) f_krige_cv(x, "rehab"))

# top picks for acute - currently, I prefer the "Exc" model
# map(c("Pen","Exc","Sph","Cir"), function(x) f_krige_cv(x, "acute"))

# top picks for acute - currently, I prefer the "Exc" model
# map(c("Sph", "Exc", "Ste", "Cir", "Lin", "Pen"), function(x) f_krige_cv(x, "rehab"))

```


```{r, include=F}
### test space for visualisaing interpolated values
vgm_model <- "Sph"

# kriged_layer <-
#   krige(
#     acute_time ~ 1, df_times, pnts,
#     model=fit.variogram(variogram(acute_time~1, df_times), model=vgm(vgm_model))
#   ) %>%
#   as.data.frame() %>%
#   rasterFromXYZ(crs=4326)

kriged_layer <-
  krige(
    rehab_time ~ 1, df_times, pnts,
    model=fit.variogram(variogram(rehab_time~1, df_times), model=vgm(vgm_model))
  ) %>%
  as.data.frame() %>%
  rasterFromXYZ(crs=4326)

bins <- c(0, 30, 60, 120, 180, 240, 300, 360, Inf)
pal <- colorBin("YlOrRd", domain = readRDS("input/layers/SA1s_acute.rds")$mean, bins = bins, na.color="transparent")

rehab_centres <- c(
  "Sunshine Coast University Hospital",
  "Central West Sub-Acute Service",
  "Gympie Hospital",
  "Rockhampton Hospital",
  "Roma Hospital"
)
acute_centres <- c(
  "Brain Injury Rehabilitation Unit",
  "Gold Coast University Hospital",
  "Townsville University Hospital"
)

df_centres <- read.csv("input/centres.csv") 
names(df_centres) <- c("centre_name", "address", "x", "y")
df_centres <- df_centres %>%
  filter(centre_name %in% c(rehab_centres, acute_centres)) %>%
  mutate(
    care_type=ifelse(centre_name %in% acute_centres, "acute", "rehab"),
    popup=paste0(
      "<b>Centre name: </b>", centre_name, "<br>",
      "<b>Care type: </b>", ifelse(care_type=="acute", "Acute care", "Rehabilitation care"), "<br>",
      "<b>Address: </b>", address, "<br>"
    )
  )
centre_icons <- iconList(
  acute=makeIcon(iconUrl = "input/imgs/acute_care2.png", iconWidth = 783/18, iconHeight = 900/18),
  rehab=makeIcon(iconUrl = "input/imgs/rehab_care.png", iconWidth = 783/18, iconHeight = 783/18)
)

leaflet() %>% 
  addSearchOSM(options=searchOptions(moveToLocation=FALSE, zoom=NULL)) %>%
  addMapPane(name = "layers", zIndex = 200) %>%
  addMapPane(name = "maplabels", zIndex = 400) %>%
  addMapPane(name = "markers", zIndex = 205) %>%
  addProviderTiles("CartoDB.VoyagerNoLabels") %>%
  addProviderTiles("CartoDB.VoyagerOnlyLabels",
                   options = leafletOptions(pane = "maplabels"),
                   group = "map labels") %>%
  addRasterImage(
    data=kriged_layer,
    x=raster(kriged_layer, layer=1),
    options=leafletOptions(pane="layers"),
    colors=pal
  ) %>%
  addMarkers(
    lng=df_centres$x, lat=df_centres$y, 
    icon=centre_icons[df_centres$care_type],
    popup=df_centres$popup,
    options=leafletOptions(pane="markers")
  )

```


# kriging - generate interpolations
```{r}
lzn_kriged_acute <- krige(acute_time ~ 1, df_times, pnts2, model=lzn_fit_acute)%>%
  as.data.frame()

krige(acute_time ~ 1, df_times, pnts, model=lzn_fit_acute)%>%
  as.data.frame() %>%
  ggplot(aes(X, Y)) + 
  geom_tile(aes(fill=var1.pred)) + 
  coord_equal()+
  scale_fill_gradient(low = "yellow", high="red")

lzn_kriged_rehab <- krige(rehab_time ~ 1, df_times, pnts2, model=lzn_fit_rehab)%>%
  as.data.frame()

krige(rehab_time ~ 1, df_times, pnts, model=lzn_fit_rehab)%>%
  as.data.frame() %>%
  ggplot(aes(X, Y)) + 
  geom_tile(aes(fill=var1.pred)) + 
  coord_equal()+
  scale_fill_gradient(low = "yellow", high="red")

# save the kriging interpolation layers that don't have the supplementary coordinates within SA1s
# These are used for the web app, whereas the 'lzn_kriged___' are used for aggregations.
krige(acute_time ~ 1, df_times, pnts, model=lzn_fit_acute) %>% 
  as.data.frame() %>%
  rasterFromXYZ(crs=4326) %>%
  saveRDS("input/layers/kriged_acute.rds")

krige(rehab_time ~ 1, df_times, pnts, model=lzn_fit_rehab) %>% 
  as.data.frame() %>%
  rasterFromXYZ(crs=4326) %>%
  saveRDS("input/layers/kriged_rehab.rds")
```

# add interpolation layer to map
```{r}
ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_bounary, color="blue") +
  labs( x = "Longitude", y = "Latitude") +
  coord_sf(xlim = c(100.00, 160.00), ylim = c(-45.00, -10.00), expand = T) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  geom_tile(data=lzn_kriged_acute, aes(X, Y, fill=var1.pred)) +
  geom_point(data=as.data.frame(df_times), aes(x,y)) +
  scale_fill_gradient(low = "yellow", high="red")

```

# Join the SA polygons together depending on level (SA1 vs SA2)
```{r}
do_union <- function(x){
  pb$tick()
  st_union(x, is_coverage = TRUE)
}

aggregate_by_SA <- function(qld_sf, SA_number, use_dissolve=FALSE){
  sa_main <- glue::glue('SA{SA_number}_{ifelse(SA_number %in% c(3,4), "CODE", "MAIN")}16')
  message(glue::glue('----- grouping polygons within SA{SA_number} -----'))
  if(use_dissolve){
    return(rmapshaper::ms_dissolve(qld_sf, sa_main))
  }
  
  num_ticks <- length(unique(qld_sf[[sa_main]]))
  pb <<- progress_bar$new(
    format = "[:bar] :current/:total (:percent) elapsed :elapsed eta :eta", 
    total=num_ticks
  )
  
  qld_sf %>%
    # https://gis.stackexchange.com/questions/321281/using-sf-to-combine-polygons-that-share-borders
    group_by(across(all_of(sa_main))) %>%
    summarize(geometry=do_union(geometry))
}

qld_SA2s <- aggregate_by_SA(qld_sf=qld_SAs, SA_number=2, use_dissolve=TRUE)
qld_SA1s <- aggregate_by_SA(qld_sf=qld_SAs, SA_number=1, use_dissolve=TRUE)

saveRDS(qld_SA2s, "input/layers/qld_SA2s.rds")
saveRDS(qld_SA1s, "input/layers/qld_SA1s.rds")

ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_SA2s, aes(fill=1)) +
  coord_sf(xlim = c(138, 153.00), ylim = c(-30.00, -10.00), expand = T) +
  guides(fill="none") +
  scale_fill_gradient(low = "yellow", high="red")
```

# find out why there is a mix of polygons and multipolygons... 
##### spoiler: multipolygons where there are islands
```{r}
mlt_idx <- c()
for(i in 1:nrow(qld_SA2s)){
  geo <- qld_SA2s$geometry[i]
  cl <- class(geo[[1]])[2]
  if(cl=="MULTIPOLYGON"){
    mlt_idx <- append(mlt_idx, i)
  }
}

ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_SA2s[mlt_idx,], aes(fill=1)) +
  geom_sf(data=qld_SA2s[-mlt_idx,], aes(fill=0)) +
  coord_sf(xlim = c(138, 153.00), ylim = c(-30.00, -10.00), expand = T) +
  guides(fill="none") +
  scale_fill_gradient(low = "yellow", high="red")
```

# spatial join from SA2 polygons to interpolated values
```{r}
get_SA_agged_times <- function(lzn_kriged_df, SA_number, save_path=NULL){
  coordinates(lzn_kriged_df) <- ~ X + Y
  lzn_kriged_sf <- st_as_sf(lzn_kriged_df)
  lzn_kriged_sf <- st_set_crs(lzn_kriged_sf, 4326)
  lzn_kriged_sf <- st_transform(lzn_kriged_sf, crs = 4326)
  
  qld_SAs <- readRDS(glue::glue("input/layers/qld_SA{SA_number}s.rds"))
  qld_SAs <- st_transform(qld_SAs, crs = 4326)
  qld_SAs_with_int_times <- st_join(qld_SAs, lzn_kriged_sf)
  
  # https://ryanpeek.org/2019-04-29-spatial-joins-in-r/
  SA_id <- glue::glue('SA{SA_number}_{ifelse(SA_number %in% c(3,4), "CODE", "MAIN")}16')
  SAs_agg_times <- 
    qld_SAs_with_int_times %>%
    na.omit() %>%
    as.data.frame() %>%
    group_by(across(all_of(SA_id))) %>%
    summarize(mean=mean(var1.pred, na.omit=TRUE))
  
  SAs_agg_times <- merge(qld_SAs, SAs_agg_times, all.x=TRUE)
  
  if(!is.null(save_path)){
    saveRDS(SAs_agg_times, save_path)
  }
  return(SAs_agg_times)
}

# write aggregated interpolation layers to disk
get_SA_agged_times(lzn_kriged_df=lzn_kriged_acute, SA_number=1, save_path="input/layers/SA1s_acute.rds")
get_SA_agged_times(lzn_kriged_df=lzn_kriged_acute, SA_number=2, save_path="input/layers/SA2s_acute.rds")
get_SA_agged_times(lzn_kriged_df=lzn_kriged_rehab, SA_number=1, save_path="input/layers/SA1s_rehab.rds")
get_SA_agged_times(lzn_kriged_df=lzn_kriged_rehab, SA_number=2, save_path="input/layers/SA2s_rehab.rds")


qld_SA2s_agged_times <- readRDS("input/layers/SA2s_acute.rds")
ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_SA2s_agged_times, aes(fill=mean), col = NA) +
  coord_sf(xlim = c(138, 153.00), ylim = c(-30.00, -10.00), expand = T) +
  guides(fill="none") +
  scale_fill_gradient(low = "yellow", high="red")

```

# spatial join from SA1 polygons to interpolated values
```{r}
qld_SA1s_agged_times <- readRDS("input/layers/SA1s_acute.rds")
ggplot(data = world) +
  geom_sf() +
  geom_sf(data=qld_SA1s_agged_times, aes(fill=mean), col = NA) +
  coord_sf(xlim = c(138, 153.00), ylim = c(-30.00, -10.00), expand = T) +
  guides(fill="none") +
  scale_fill_gradient(low = "yellow", high="red")

```

# test space for leaflet
```{r}
qld_SA2s_agged_times <- readRDS("input/layers/SA2s_acute.rds")
bins <- c(0, 30, 60, 120, 180, 240, 300, 360, Inf)
pal <- colorBin("YlOrRd", domain = qld_SA2s_agged_times$mean, bins = bins)

df_locations <- read.csv("input/QLD_locations_with_RSQ_times_20220210.csv")
df_locations <- mutate(df_locations, popup=paste0(
  "<b>Location: </b>", location, "<br>",
  "<b>Acute care destination: </b>", acute_care_centre, "<br>",
  "<b>Time to acute care (minutes): </b>", acute_time, "<br>",
  "<b>Rehab care destination: </b>", rehab_centre, "<br>",
  "<b>Time to rehab care (minutes): </b>", rehab_time, "<br>"
))

rehab_centres <- c(
  "Sunshine Coast University Hospital",
  "Central West Sub-Acute Service",
  "Gympie Hospital",
  "Rockhampton Hospital",
  "Roma Hospital"
)
acute_centres <- c(
  "Brain Injury Rehabilitation Unit",
  "Gold Coast University Hospital",
  "Townsville University Hospital"
)

df_centres <- read.csv("input/centres.csv") 
names(df_centres) <- c("centre_name", "address", "x", "y")

df_centres <- df_centres %>%
  # mutate(centre_name = str_trim(centre_name))%>%
  filter(centre_name %in% c(rehab_centres, acute_centres)) %>%
  mutate(
    care_type=ifelse(centre_name %in% acute_centres, "acute", "rehab"),
    popup=paste0(
      "<b>Centre name: </b>", centre_name, "<br>",
      "<b>Care type: </b>", ifelse(care_type=="acute", "Acute care", "Rehabilitation care"), "<br>",
      "<b>Address: </b>", address, "<br>"
    )
  )

centre_icons <- iconList(
  acute=makeIcon(iconUrl = "input/imgs/acute_care2.png", iconWidth = 783/18, iconHeight = 900/18),
  rehab=makeIcon(iconUrl = "input/imgs/rehab_care.png", iconWidth = 783/18, iconHeight = 900/18)
)
leaflet() %>% 
  addSearchOSM(options=searchOptions(moveToLocation=FALSE, zoom=NULL)) %>%
  addMapPane(name = "polygons", zIndex = 200) %>%
  addMapPane(name = "maplabels", zIndex = 2102) %>%
  addProviderTiles("CartoDB.VoyagerNoLabels") %>%
  addProviderTiles("CartoDB.VoyagerOnlyLabels",
                   options = leafletOptions(pane = "maplabels"),
                   group = "map labels") %>%
  addPolygons(
    data=qld_SA2s_agged_times, 
    fillColor = ~pal(mean), 
    color="black", 
    fillOpacity=1,
    weight=1,
    group="polys",
    options = leafletOptions(pane = "polygons")
  ) %>%
  addLayersControl(
    baseGroups = "CartoDB.PositronNoLabels",
    overlayGroups = c("map labels","polys")
  ) %>%
  addCircleMarkers(
    lng=df_locations$x, lat=df_locations$y, 
    radius=2, fillOpacity=0,
    popup=df_locations$popup
  ) %>%
  addMarkers(
    lng=df_centres$x, lat=df_centres$y, 
    icon=centre_icons[df_centres$care_type],
    popup=df_centres$popup
  )
```


```{r}

x <- readRDS("input/layers/kriged_acute.rds")
SA1s_acute <- readRDS("input/layers/SA1s_acute.rds")
roads_sf <- st_read("input/baseline_roads_and_tracks/Baseline_roads_and_tracks.shp")

bins <- c(0, 30, 60, 120, 180, 240, 300, 360, Inf)
pal <- colorBin("YlOrRd", domain = SA1s_acute$mean, bins = bins, na.color="transparent")
rasterPal <- colorBin("YlOrRd", domain = x$var1.pred, bins = bins, na.color="transparent")

leaflet() %>% 
  addSearchOSM(options=searchOptions(moveToLocation=FALSE, zoom=NULL)) %>%
  addMapPane(name = "layers", zIndex = 200) %>%
  addMapPane(name = "maplabels", zIndex = 210) %>%
  addProviderTiles("CartoDB.VoyagerNoLabels") %>%
  addProviderTiles("CartoDB.VoyagerOnlyLabels",
                   options = leafletOptions(pane = "maplabels"),
                   group = "map labels") %>%
  # addPolygons(
  #         data=qld_SA1s_agged_times,
  #         fillColor = ~pal(mean),
  #         color="black",
  #         fillOpacity=1,
  #         weight=1,
  #         options = leafletOptions(pane = "layers")
  #       ) %>%
  addRasterImage(
    data=x,
    x=raster(x, layer=1),
    options=leafletOptions(pane="layers"),
    colors=pal
  ) %>%
  addPolylines(data=roads_sf[roads_sf$ROADTYPE==2,], 
               weight=1, opacity=1, fillOpacity=1, color="black",
               options=leafletOptions(pane="maplabels"))

ggplot() + geom_sf(data=roads_sf[roads_sf$ROADTYPE==3,])
```

```{r, include=F}
# write aggregate data to disk so that it can be included in app
SA2s_acute <- readRDS("input/layers/SA2s_acute.rds")
SA2s_rehab <- readRDS("input/layers/SA2s_rehab.rds")
SA1s_acute <- readRDS("input/layers/SA1s_acute.rds")
SA1s_rehab <- readRDS("input/layers/SA1s_rehab.rds")
qld_SAs <- st_read("input/qld_sa_zones/MB_2016_QLD.shp")
SA2s_data <- 
  as.data.frame(qld_SAs) %>%
  dplyr::select(SA2_NAME16, SA2_MAIN16) %>%
  distinct() %>%
  left_join(., dplyr::select(as.data.frame(SA2s_acute), SA2_MAIN16, acute_time=mean)) %>%
  left_join(., dplyr::select(as.data.frame(SA2s_rehab), SA2_MAIN16, rehab_time=mean)) %>%
  na.omit()

SA1s_data <- 
  as.data.frame(qld_SAs) %>%
  dplyr::select(SA1_MAIN16) %>%
  distinct() %>%
  left_join(., dplyr::select(as.data.frame(SA1s_acute), SA1_MAIN16, acute_time=mean)) %>%
  left_join(., dplyr::select(as.data.frame(SA1s_rehab), SA1_MAIN16, rehab_time=mean)) %>%
  na.omit()

write.csv(SA2s_data, "input/SA2s_data.csv", row.names = FALSE)
write.csv(SA1s_data, "input/SA1s_data.csv", row.names = FALSE)

```
